Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-08 windheat.tex

For a stellar population that forms impulsively, the rate of mass and energy injection per unit stellar mass at age $t$ are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to
contributions from main sequence (MS) stars and post-main sequence
(PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is the efficiency
with which main sequence winds thermalize their energy with the rest
of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$
is the IMF, which we take to be of the Salpeter form $\mu|_{\Mstar}
\propto M_{\star}^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low
mass end and at $100 \Msun$ on the high mass end, where $\bar{m}_*$ =
0.35 $\Msun$ is the corresponding mean stellar mass.  Our optimistic
assumption of 100 per cent thermalization efficiency of stellar winds
is at least partially offset by our conservative assumption to neglect
energy input from core collapse supernovae, which contribute a
comparable energy injection rate to massive stellar winds.

The quantity of mass lost in PMS winds $\Delta M(t)$ is estimated from
the expression given by \citet{CiottiOstriker:2007a} (their eq.~[10]),
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun,
\end{cases}
\end{align}
where $M_{\rm TO}$ is the turn-off mass, which at time $t < t_{\rm h}$ is calculated as
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$.  This functional fit is designed
to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9)
for massive stars while asymptoting to the formula provided by
\citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late
times ($t \gsim 10^8$ years).

@@

\subsection{Over Star Formation History}
Generalizing to an arbitrary star formation history $S(t)$, the total rate of mass and energy input can be written as
\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1 \label{eq:MDotSFH}\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1, \label{eq:EDotSFH}
\end{align}
resulting in a wind heating parameter of 
\begin{align}
  v_w^2(t) &=2 \dot{E}(t)/\dot{M}(t).
\end{align}
The mass return parameter will be
\begin{align}
\eta = \frac{\dot{M}(t)}{\int_0^t S(t_1) {\rm d}t_1} \th
\end{align}

>>>>

In the case of a single implusive burst of star formation, the mass and energy injection rate per unit stellar mass at time $t$ after the burst are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to
contributions from main sequence (MS) stars and post-main sequence
(PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is the efficiency
with which main sequence winds thermalize their energy with the rest
of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$
is the IMF, which we take to be of the Salpeter form $\mu|_{\Mstar}
\propto M_{\star}^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low
mass end and at $100 \Msun$ on the high mass end, where $\bar{m}_*$ =
0.35 $\Msun$ is the corresponding mean stellar mass.  Our optimistic
assumption of 100 per cent thermalization efficiency of stellar winds
is at least partially offset by our conservative assumption to neglect
energy input from core collapse supernovae, which contribute a
comparable energy injection rate to massive stellar winds.

The quantity of mass lost in PMS winds $\Delta M(t)$ is estimated from
the expression given by \citet{CiottiOstriker:2007a} (their eq.~[10]),
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun,
\end{cases}
\end{align}
where $M_{\rm TO}$ is the turn-off mass, which at time $t < t_{\rm h}$ is calculated as
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm yr})$.  This functional fit is designed
to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9)
for massive stars while asymptoting to the formula provided by
\citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late
times ($t \gsim 10^8$ years).

@@

\subsection{Over Star Formation History}
Generalizing to an arbitrary SFH $S(t)$, the total rate of mass and energy input can be written as
\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1 \label{eq:MDotSFH}\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1, \label{eq:EDotSFH}
\end{align}
resulting in a wind heating parameter of 
\begin{align}
  v_w^2(t) &=2 \dot{E}(t)/\dot{M}(t).
\end{align}
The mass return parameter will be
\begin{align}
\eta = \frac{\dot{M}(t)}{\int_0^t S(t_1) {\rm d}t_1} \th
\end{align}
